---
title: "Network Analytics Individual Assignment"
author: "Jessica Gumba"
date: '2022-10-18'
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, warn=FALSE, message=FALSE} 
library("igraph")
library("ggplot2")
library("dplyr")

# Let's take our text file of space-separated edges and put it into an iGraph graph
subset1000 <- read.graph(file="data/graph_subset_rank1000.txt", format ="ncol")
subset1000cc <- read.graph(file="data/graph_subset_rank1000_cc.txt", format ="ncol")
all_data <- read.graph(file="data/graph_complete.txt", format ="ncol")
idtotitles <- read.table(file="data/id_to_titles.txt", sep = " ", quote = "\"",header = TRUE)
```

## 1. Network Structure Visualization
### (a) Let's analyse the subset file containing edges between top-ranked products.
```{r analyse}
subset1000

# How many nodes?
vcount(subset1000)

# How many edges?
ecount(subset1000)

# Get nodes
V(subset1000)

# Get edges
E(subset1000)

# Is the graph directed?
is.directed(subset1000)
```

### (b). Plot the network using the information we have on the 1000 top-ranked products.

#### Using layout: ` layout.kamada.kawai`
```{r plot_subset1000a}
plot.igraph(subset1000, 
            layout=layout.kamada.kawai, 
            vertex.label=NA, 
            vertex.size=4,
            mode = "undirected",
            main = "Network Structure - Sales Rank between 0 and 1,000",
            sub = "layout: layout.kamada.kawai")
```

#### Using default layout
```{r plot_subset1000b}
plot.igraph(subset1000, 
            vertex.label=NA, 
            vertex.size=4,
            layout=layout.auto, 
            mode = "undirected",
            main = "Network Structure - Sales Rank between 0 and 1,000",
            sub = "default layout: layout.auto")
```

#### Findings:
* From both visualizations, we can see that there are some small clusters formed.
* However, even more noticeable is a very large cluster:
  * In the kamada-kawai layout, we can see this at the bottom of the visualization.
  * In the default layout, the middle section stands out because of the large connected network, unlike the clusters surrounding it.

### Let's zoom into the largest connected component in the above network.

#### Using layout: ` layout.kamada.kawai`
```{r plot_subset1000cca}
plot.igraph(subset1000cc, 
            layout=layout.kamada.kawai, 
            vertex.label=NA, 
            vertex.size=4,
            mode = "undirected",
            main = "Sales Rank between 0 and 1,000 (Largest Component)",
            sub = "layout: layout.kamada.kawai")
```

#### Using default layout
```{r plot_subset1000ccb}
plot.igraph(subset1000cc, 
            vertex.label=NA, 
            vertex.size=4,
            layout=layout.auto, 
            mode = "undirected",
            main = "Sales Rank between 0 and 1,000 (Largest Component)",
            sub = "default layout: layout.auto")
```

## 2. Data Analysis

### (a) Out-degree distribution
* First, we will calculate the out-degree for each product. We will use the table command in R, which requires storing the full dataset into dataframe `freq_out`.
* We will use `id to titles.csv.` as our source of truth for the full list of all titles. We expect some of the titles to not have any record in the `graph_complete.csv` file; for such cases, we'll set `freq_out$freq == 0`.


```{r outdegree_calculations}
# Convert to a data frame
data_full <- as.data.frame(read.table(file = "data/graph_complete.txt"))
colnames(data_full) <- c("from","to")

# Table frequencies
freq_out <- as.data.frame(table(data_full$from))
colnames(freq_out) <- c("id","freq")

data_out<-merge(freq_out, idtotitles, by.x = "id", by.y="id", all.y=TRUE)
data_out<-data_out %>% mutate(freq = ifelse(is.na(freq), 0, freq))
data_out = data_out %>% select(-c("title"))
colnames(data_out) <- c("id","freq")

ggplot(data_out, aes(x=freq)) + 
  geom_histogram(binwidth=1, color="white", fill="#4cbea3") +
  ggtitle("Out-degree distribution") +
  labs(x="No. of Outlinks", y = "No. of Products")
```

#### Findings:
* From the histogram generated, we can conclude that the maximum number of outbound links is 5.
* Most products have number of outbound links = 4, whereas the minority have no outbound links.
* Overall, this visualization seems to display a normal distribution skewed to the left.

### (b) In-degree distribution
* Similar to above, we will use the table command in R to calculate the in-degree distribution for the same dataset, which requires storing the full dataset into dataframe `freq_in`.
* We will also use `id to titles.csv.` as our source of truth for the full list of all titles. We expect some of the titles to not have any record in the `graph_complete.csv` file; for such cases, we'll set `freq_in$freq == 0`.

```{r indegree_calculations}
# Plot distribution
freq_in <- as.data.frame(table(data_full$to))
colnames(freq_in) <- c("id","freq")

data_in<-merge(freq_in, idtotitles, by.x = "id", by.y="id", all.y=TRUE)
data_in<-data_in %>% mutate(freq = ifelse(is.na(freq), 0, freq))
#data_in = data_in %>% select(-c("title"))
colnames(data_in) <- c("id","freq","title")

ggplot(data_in, aes(x=freq)) + 
  geom_histogram(binwidth=4, fill="#4cbea3", color="black") +
  ggtitle("In-degree distribution") +
  labs(x="No. of Incoming Links", y = "No. of Products") 
```

#### Findings
* The in-degree distribution is very different from that of the out-degree. 
* It seems to follow the power-law distribution, where a vast majority of the titles do not have any incoming links (i.e. freq==0) whereas a small percentage have as much as 100+ to 500+ (freq >= 100).

### (c) In-degree distribution (log)
* Next, we will alter the previous plot such that the x-axis is the `log(freq)`.
* We will replace all `freq==0` with `freq==0.1` to avoid any error in the transformation.


#### Plot 1: Using a scatterplot on the aggreated dataframe
* We've aggregated the number of nodes per frequency (or number of incoming links) and created a scatterplot, and the results are as shown below.
* Findings are consistent with the above.

```{r log_plot}
data_in<-data_in %>% mutate(freq = ifelse(freq==0, 0.1, freq))
data_in$log<-log(as.numeric(data_in$freq))

logplot <- as.data.frame(table(as.numeric(data_in$freq)))
logplot$log<-log(as.numeric(paste(logplot$Var1)))
colnames(logplot) <- c("inlinks","freq","log")

#Convert
ggplot(data=logplot, aes(x=log, y=freq)) +
  geom_point() +
  ggtitle("Aggregate View - In-degree Distribution") +
  labs(x="Log no. of Incoming Links", y = "No. of Nodes")
```


#### Plot 2: Using a histogram
* The histogram results are consistent with the findings above. Majority of the titles fall under log(0.1) = -2.3.
* We can also more clearly see the distribution:
  * Many titles have at least 1 incoming link, at log(1) = 0.
  * Very few have a high number of incoming links, evident in the long tail of the distribution.
  
  
```{r}
ggplot(data_in, aes(x=log)) + 
  geom_histogram(binwidth=0.25, fill="#4cbea3", color="black") +
  ggtitle("In-degree distribution") +
  labs(x="Log no. of Incoming links", y = "No. of Products") + theme(panel.background = element_blank())
```

## 3. Computing for the Mean, Standard Deviation, Maximum

#### Findings
* As expected from the power law distribution generated above, the average number of incoming links hovers around 3-4, whereas the highest value is at 549 which is far above the mean. This further proves the big gap between the number of incoming links in our dataset.
* The standard deviation is quite high at ~6, showing a high amount of variation & dispersion in the number of incoming links in our dataset.

```{r}
mean(data_in$freq)
sd(data_in$freq)
max(data_in$freq)
```
## 4. Top 10 Products with the Highest Inbound Co-purchase Links

```{r books}
top10<-data_in[order(-data_in$freq),][1:10,c("id", "title", "freq")]
top10$title
```

